---
title: "Chapter 2"
author: "Jorge Cimentada"
date: "5/16/2019"
output: html_document
---

## Quiz answers

- Given the following data frame, how do I create a new column called “3” that contains the sum of 1 and 2? You may only use $, not [[. What makes 1, 2, and 3 challenging as variable names?

```{r}
df <- data.frame(runif(3), runif(3))
names(df) <- c(1, 2)
```

Solution: 

```{r}
df$`3` <- df$`1` + df$`2`
```

- In the following code, how much memory does y occupy?

```{r}
x <- runif(1e6)
y <- list(x, x, x)
```

Not entirely sure. I though it was the same as `x` because R shared the same memory when an object is repeated. However, it is not either that 3 times the size of `x`.

For example, the size of `x` is

```{r}
object.size(x)
```

That size multiplied by three is

```{r}
(object.size(x) * 3)
```

and the size of `y` is

```{r}
object.size(y)
```

There seems to be a small diference which I am not sure where it comes from.


- On which line does `a` get copied in the following example?

```{r}
a <- c(1, 5, 3, 2)
b <- a
b[[1]] <- 10
```

I'm not entirely sure. I know that for assignments (`<-`) there is an intermediate copy being made in the first line but I'm not sure whether the refer to this one. However, I do know that in the second line `b` and `a` are the same object.

```{r}
a <- c(1, 5, 3, 2)
b <- a

.Internal(inspect(a))
.Internal(inspect(b))
```
Both have the same memory signature, thus occupying the same space. Moreover, once `b` is different, the memory signature changes.

```{r}
b[[1]] <- 10
.Internal(inspect(a))
.Internal(inspect(b))
```

Now how does this translate to memory management in R? More specifically, am I occupying more memory by having the same object with two different names?

```{r}
lobstr::mem_used()
pryr::mem_change(a <- 1:10)
pryr::mem_change(b <- a)
lobstr::mem_used()
```
It seems like it does! However, I presume this might be extra space only allocated to the new name because it's much smaller than the increase of assigning `a`. If `a` would be copied all over again, then I assume we would see the same increas in memory.

```{r}
library(lobstr)
set.seed(1014)
```

Exercises 2.2.2

- Explain the relationship between a, b, c and d in the following code:

```{r}
a <- 1:10
b <- a
c <- b
d <- 1:10
```

`a`, `b` and `c` share the same memory address and `d` has a different one.

```{r}
objs <- list(a, b, c, d)
obj_addrs(objs)
```

However, for `d`, it's a bit different. They share different memory addresses.

```{r}
obj_addr(a)
obj_addr(d)
```

That would think that since the memory address is different, then the memory is duplicated whenever you create the same two objects with different names:

```{r}
rm(list = ls())
pryr::mem_change(y <- 1:10)
obj_addr(y)
pryr::mem_change(z <- 1:10)
obj_addr(z)
```

This is still a pending question asked [here](https://twitter.com/cimentadaj/status/1129047253346332674)


```{r}
pryr::mem_used()
pryr::mem_change(a <- 1:10)
pryr::mem_change(b <- a)
pryr::mem_used()
```

