---
title: "Chapter 2"
author: "Jorge Cimentada"
date: "5/16/2019"
output: html_document
---

## Quiz answers

```{r }
library(lobstr)
```

- Given the following data frame, how do I create a new column called “3” that contains the sum of 1 and 2? You may only use $, not [[. What makes 1, 2, and 3 challenging as variable names?

```{r}
df <- data.frame(runif(3), runif(3))
names(df) <- c(1, 2)
```

Solution: 

```{r}
df$`3` <- df$`1` + df$`2`
```

- In the following code, how much memory does y occupy?

```{r}
x <- runif(1e6)
y <- list(x, x, x)
```

Not entirely sure. I though it was the same as `x` because R shared the same memory when an object is repeated. However, it is not either that 3 times the size of `x`.

For example, the size of `x` is

```{r}
obj_size(x)
```

That size multiplied by three is

```{r}
(obj_size(x) * 3)
```

and the size of `y` is

```{r}
obj_size(y)
```

There seems to be a small diference which I am not sure where it comes from. I think the difference comes from the space occupied by the list function.


- On which line does `a` get copied in the following example?

```{r}
a <- c(1, 5, 3, 2)
b <- a
b[[1]] <- 10
```

I'm not entirely sure. I know that for assignments (`<-`) there is an intermediate copy being made in the first line but I'm not sure whether to refer to this one. However, I do know that in the second line `b` and `a` are the same object.

```{r}
a <- c(1, 5, 3, 2)
b <- a

.Internal(inspect(a))
.Internal(inspect(b))
```

Both have the same memory signature, thus occupying the same space. Moreover, once `b` is different, the memory signature changes.

```{r}
b[[1]] <- 10
.Internal(inspect(a))
.Internal(inspect(b))
```

Now how does this translate to memory management in R? More specifically, am I occupying more memory by having the same object with two different names?

```{r}
lobstr::mem_used()
pryr::mem_change(a <- 1:10)
pryr::mem_change(b <- a)
lobstr::mem_used()
```

It seems like it does! However, I presume this might be extra space only allocated to the new name because it's much smaller than the increase of assigning `a`. If `a` would be copied all over again, then I assume we would see the same increase in memory.

```{r}
library(lobstr)
set.seed(1014)
```

Exercises 2.2.2

1. Explain the relationship between a, b, c and d in the following code:

```{r}

a <- 1:10
b <- a
c <- b
d <- 1:10

```

`a`, `b` and `c` share the same memory address and `d` has a different one.

```{r}
objs <- list(a, b, c, d)
obj_addrs(objs)
```

However, for `d`, it's a bit different. They share different memory addresses.

```{r}
obj_addr(a)
obj_addr(d)
```
That would think that since the memory address is different, then the memory is duplicated whenever you create the same two objects with different names:

```{r}

rm(list = ls())
pryr::mem_change(y <- 1:10)
obj_addr(y)
pryr::mem_change(z <- 1:10)
obj_addr(z)

```


```{r}
# This should return the same address but increase in memory due to the name
# binding.

# First create the environment hash table for all name bindings **and** `a`
# The total size is `a` + hash table creation, which is why it's so big
# compared to the increases below.
pryr::mem_change(a <- 1:10)

# Then bind `a` another name `b`. This should have the same memory address
# as `a`. The increase in memory should be lower than the operations below
# because it's pointing to the same memory address.
pryr::mem_change(b <- a)
# If run, adds more memory, so commenting for now
# lobstr::obj_addr(a) == lobstr::obj_addr(b)

# m should have a different address from `a` and `b`
# The increase in 936 B is probably due to the space occupied by the
# name and the values 1:10
pryr::mem_change(m <- 1:10)
# If run, adds more memory, so commenting for now
# lobstr::obj_addr(a) == lobstr::obj_addr(m)

# That is why running the same below will always add the same thing
pryr::mem_change(p <- 1:10)
pryr::mem_change(u <- 1:10)
pryr::mem_change(i <- 1:10)

```

The question asked [here](https://twitter.com/cimentadaj/status/1129047253346332674) helped to achieve that.

Exercises 2.2.2

2. The following code accesses the mean function in multiple ways. Do they all point to the same underlying function object? Verify this with lobstr::obj_addr().

```{r }
all_means <- list(
  mean,
  base::mean,
  get("mean"),
  evalq(mean),
  match.fun("mean")
)

all_adr <- lobstr::obj_addrs(all_means)

all(all_adr[1] == all_adr)
```

3. By default, base R data import functions, like read.csv(), will automatically convert non-syntactic names to syntactic ones. Why might this be problematic? What option allows you to suppress this behaviour?

You can set `make.names` equal to `FALSE` to avoid this problem. One case that I can think of for why `read.csv` shouldn't convert to syntactic is that this varies by the locale of the computer. Meaning that you might get different results on different computers if `make.names` is applied. Another problem is that `make.names` converts non-supported characters to `.`. For example, if you had a column called `@` with the email of people, make names with change that to `X.`:

```{r }
make.names("@")
```

4. What rules does make.names() use to convert non-syntactic names into syntactic ones?

If it starts with a dot followed by a number or a underscore, the character "X" is prepended.

```{r }
make.names(".2")
make.names("_.2")
```

All invalid characters are translated to ".". 

```{r }
make.names("random_name_with_@")
```

A missing value is translated to "NA".  Names which match R keywords have a dot appended to them.

```{r }
make.names("if")
```

Duplicated values are altered by make.unique.

```{r }
make.names(c("c", "c"), unique = TRUE)
```

5. I slightly simplified the rules that govern syntactic names. Why is .123e1 not a syntactic name? Read ?make.names for the full details.

Because it beings with a `.` followed by a number. `make.names` forbids that. Quoting from `make.names`:

- If it starts with a dot followed by a number or a underscore, the character "X" is prepended.


# Notes

```{r }

# If you define an x
x <- 1:3
# Here's its memory address
cat(tracemem(x), "\n")

# If you modify it, it creates a copy
x[[1]] <- 2L

# If you bind it to another name, it doesn't create a copy
# but just points the `y` name to the memory address we just
# saw.
y <- x

# However, the moment we modify it, it creates another copy
y[[3]] <- 4L
# But further modifications don't copy
y[[2]] <- 5L
y[[1]] <- 1L

# The same thing with `z`
# Doesn't create copy
z <- y
# Creates copy after first modification
z[[1]] <- 3L
# Doesn't modify
z[[1]] <- 9L
z[[3]] <- 4L


# For lists, it's slight different because it creates
# "shallow" copies, reciclyng memory addresses when
# possible.
l1 <- list(1, 2, 3)

# No copy made
l2 <- l1
# Copy made but only for the third vector
l2[[3]] <- 3L

ref(l1, l2)
```

2.3.6 Exercises

1. Why is tracemem(1:10) not useful?

Because it doesn't bind a name to that memory address meaning that you cannot check where the copies are being made. Think of it as occupying a memory address to a millisecond and then it's freed. An example:

```{r }
# It increases after 1:10 but decreases in the third `mem_used` call
mem_used(); 1:10; mem_used(); mem_used()
```

2. Explain why tracemem() shows two copies when you run this code. Hint: carefully look at the difference between this code and the code shown earlier in the section.

```{r }
x <- c(1L, 2L, 3L)
tracemem(x)

x[[3]] <- 4
```

It creates two copies because of modification of value and because of of coercing from an integer vector to a double vector. We could fix the above by just specifying `4` as an integer `4L`.

3. Sketch out the relationship between the following objects:

```{r }
a <- 1:10
b <- list(a, a)
c <- list(b, a, 1:10)
```

- `b` is a list that contains the same two memory addresses belonging to `a`:
```{r}
ref(b)
```

- `c` is a list that contains the memory address to the list `b` (including the shared memory addresses of `a`), the same memory address for `a` and a new memory address for `1:10`:

```{r}
ref(c)
``` 

4. What happens when you run this code?

```{r }
x <- list(1:10)
x[[2]] <- x
```

Draw a picture.

- Creates a memory address for `list`, a memory address for `1:10` and binds it to x.
- Create a copy of the list `x` and put it in the 2nd slot.

This means that the memory address of `list` is copied but everything else is shared:

```{r }
ref(x)
```

It's the same as this:

```{r }
tracemem(x)
y <- x
y[[2]] <- x[[2]]

ref(y)
```


## Notes

Does repeating the same number N times makes the size N times bigger?

```{r }
# character
a <- c("a")
obj_size(a)
# For character no
obj_size(rep(a, 100))

# double
a <- 1
obj_size(a)
# For double no
obj_size(rep(a, 100))


# integer
a <- 1L
obj_size(a)
# For integer even less. Is this because it recycles
# decimals?
obj_size(rep(a, 100))



a <- list(1, 2, 3)
obj_size(a)

b <- rep(a, 3)
obj_size(b)

ref(b)

```


2.4.1 Exercises

1. In the following example, why are object.size(y) and obj_size(y) so radically different? Consult the documentation of object.size().

```{r }
tst <- list(runif(1e4))
y <- rep(tst, 100)

object.size(y)
#> 8005648 bytes
obj_size(y)
#> 80,896 B
```

Simply because `object.size` doesn't take into account that a list has shared references and thus the size doesn't increase linearly.

2. Take the following list. Why is its size somewhat misleading?

```{r}
funs <- list(mean, sd, var)
obj_size(funs)
#> 17,608 B
```

3. Predict the output of the following code:

```{r }
a <- runif(1e6)
# Couldn't do it
obj_size(a)

b <- list(a, a)
# Slightly more due to a list of length two
obj_size(b)
# Check the difference
(obj_size(b) - obj_size(a)) == obj_size(list(NULL, NULL))

# The same because everything is being reference to: nothing
# is being created between a and b
obj_size(a, b)

# Here you're changing the first value of the vector in the first
# slot of `b`. Maybe you think that changing only a single value
# will not have an effect, but R makes a complete copy of the first
# vector simply because of changing this number.
b[[1]][[1]] <- 10
# Double the size because of copy-on-modification
obj_size(b)
obj_size(a, b)

# Creates another copy and thus the size increases again
# by around 8B.
b[[2]][[1]] <- 10
obj_size(b)
obj_size(a, b)
```


## Notes

```{r }

a <- c(1, 2, 3)
obj_addr(a)

a[[3]] <- 4
obj_addr(a)

x <- data.frame(matrix(runif(5 * 1e4), ncol = 5))
medians <- vapply(x, median, numeric(1))

for (i in seq_along(medians)) {
  x[[i]] <- x[[i]] - medians[[i]]
}

tracemem(x)
for (i in seq_along(medians)) {
  x[[i]] <- x[[i]] - medians[[i]]
}

y <- as.list(x)
tracemem(y)
for (i in seq_along(medians)) {
  y[[i]] <- y[[i]] - medians[[i]]
}


e1 <- rlang::env(a = 1, b = 2, c = 3)
e2 <- e1

e1$c <- 4

```

2.5.3 Exercises

1. Explain why the following code doesn’t create a circular list.

```{r }
x <- list()
x[[1]] <- x
```

Because since `x` is being modified, it creates a copy of `x`, effectively changing it's memory address. For example:

```{r}
x <- list()
tracemem(x)
x[[1]] <- x
```

When you run `ref`, you can see that the initial memory address of `x` is now in the first slot of the list while the list itself was copied and has a different memory address:

```{r}
ref(x)
```

2. Wrap the two methods for subtracting medians into two functions, then use the ‘bench’ package (Hester 2018) to carefully compare their speeds. How does performance change as the number of columns increase?

```{r }
slow_m <- function(dt) {
  all_medians <- vapply(dt, median, numeric(1))
  for (i in seq_along(all_medians)) {
    dt[[i]] <- dt[[i]] - all_medians[i]
  }
  dt
}

fast_m <- function(dt) {
  all_medians <- vapply(dt, median, numeric(1))
  for (i in seq_along(all_medians)) {
    dt[[i]] <- dt[[i]] - all_medians[i]
  }
  as.data.frame(dt)
}

res <- bench::press(
  col = seq(5, 2000, 50),
  {
    dt <- data.frame(matrix(runif(5 * 1e6), ncol = col))
    dl <- as.list(dt)
    bench::mark(
      min_iterations = 5,
      slow_m(dt),
      fast_m(dl)
    )
  }
)

res %>%
  mutate(expression = as.character(expression)) %>% 
  ggplot(aes(col, median, color = expression)) +
  geom_smooth()
```

3. What happens if you attempt to use tracemem() on an environment?

```{r }
tracemem(rlang::env())
```

`tracemem` is only useful for things which are copied and modified, since that would allow us to look at the memory trace. As the `tracemem` docs says:

> It is not possible to trace functions, as this would conflict with trace and it is not useful to trace NULL, environments, promises, weak references, or external pointer objects, as these are not duplicated.


