---
title: 'Chapter 3 answers'
author: Jorge Cimentada
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../figs/",
                      fig.align = "center",
                      fig.asp = 0.618,
                      out.width = "80%")
```


- 3.2.5 Exercises

1. How do you create raw and complex scalars? (See ?raw and ?complex.)

You can create raw vectors with:

```{r}
# create from scratch (empty, 10 is the length)
raw(10)

# or convert from existing vectors
as.raw(c(99L, 2L, 10L))
```

You can create complex vectors:

```{r }
z <- complex(real = stats::rnorm(100), imaginary = stats::rnorm(100))
z
```

2. Test your knowledge of the vector coercion rules by predicting the output of the following uses of c():

```{r }
# Output: numeric
c(1, FALSE)

# Output: character
c("a", 1)

# Output: numeric
c(TRUE, 1L)
```

3. Why is 1 == "1" true? Why is -1 < FALSE true? Why is "one" < 2 false?

```{r}
# Why is this true?
# because as.numeric("1") is 1
1 == "1"

# Because FALSE is zero
-1 < FALSE

#  MISSING THISSS
"one" < 2

```


4. Why is the default missing value, NA, a logical vector? What’s special about logical vectors? (Hint: think about c(FALSE, NA_character_).)

5. Precisely what do is.atomic(), is.numeric(), and is.vector() test for?
